Software Architectures for Integrating Workflow Engines in Science Gateways

### 2016-05-13 paper.tex
done 3.2

@@

\paragraph{Integration complexity} This architecture does not require any
component in addition to the science gateway and infrastructure
(\texttt{I$_1$=2}). It involves 5 interactions: \texttt{a$_1$},
\texttt{a$_2$}, \texttt{b}, \texttt{c$_1$} and \texttt{c$_2$}
(\texttt{I$_2$=5}).

@@

\paragraph{Extensibility} Integrating a new type of workflow engine
requires to modify the science gateway as well as interactions
\texttt{b} and \texttt{c$_2$} (\texttt{E$_1$}=3). Updating a workflow
engine version requires modifications in the science gateway
(\texttt{E$_2$}=1).  Inserting a new workflow is done through
interactions \texttt{a$_1$} and \texttt{a$_2$}
(\texttt{E$_3$}=2). Adding a new infrastructure generates updates in
interactions \texttt{a$_1$}, \texttt{b}, \texttt{c$_1$} and
\texttt{c$_2$} (\texttt{E$_4$}=4).

\paragraph{Scalability} Adding a new engine instance requires a new
instance of the science gateway, which is in general not possible
(\texttt{S$_1$}=2). Specific IT setups such as load-balancing between
web server instances might be used to deal with this
issue. \closedanswerednote{Tristan}{What-about load-balancing between web
  server instances?}{Marc-e}{Would that really be considered a real
  architectural feature or an IT hack?}{Tristan}{I agree, we should
  keep IT setups out of the picture. Depending on the internals of the
  science gateway, such hacks may or may not be feasible. I added a
  sentence to mention this possibility, without changing the score.}
Distributed engines are not available by default
(\texttt{S$_2$}=1). The scheduling of tasks on the infrastructure is
as complex as in any other architecture since the workflow engine
might implement any kind of scheduling policy (\texttt{S$_3$}=0).

\paragraph{Specific features} Meta-workflows are not supported by default
(\texttt{O$_1$}=1).  Debugging is not an issue since the science
gateway can retrieve any information from the workflow engine directly
(\texttt{O$_2$}=0).

@@

>>>>

\note{sil}{since we are dreaming, it seems to me that keeping provenance properly is also something to consider. the architecture can hide some details, and make it nearly impossible to keep provenance without hard coded hacks across the layers/components - i think}

@@

\paragraph{Integration complexity} This architecture does not require any
component in addition to the science gateway and infrastructure
(\texttt{I$_1$=2}) and it involves 5 interactions: \texttt{a$_1$},
\texttt{a$_2$}, \texttt{b}, \texttt{c$_1$} and \texttt{c$_2$}
(\texttt{I$_2$=5}).

@@

\paragraph{Extensibility} Integrating a new type of workflow engine
requires modifying the science gateway as well as interactions
\texttt{b} and \texttt{c$_2$} (\texttt{E$_1$}=3). Updating a workflow
engine version requires modifications in the science gateway
(\texttt{E$_2$}=1).  Inserting a new workflow is done through
interactions \texttt{a$_1$} and \texttt{a$_2$}
(\texttt{E$_3$}=2). Adding a new infrastructure generates updates in
interactions \texttt{a$_1$}, \texttt{b}, \texttt{c$_1$} and
\texttt{c$_2$} (\texttt{E$_4$}=4).

\paragraph{Scalability} Adding a new engine instance requires a new
instance of the complete science gateway
%, which is in general not possible
\note{sil}{why not? i agree that S=2, but i do not agree that it would be impossible. you say how below}
(\texttt{S$_1$}=2). 
Specific IT setups such as load-balancing between
web server instances might be used to create new gateway instances, distributing the users, but would not solve the problem for the workload of a single workflow or user.
\closedanswerednote{Tristan}{What-about load-balancing between web
  server instances?}{Marc-e}{Would that really be considered a real
  architectural feature or an IT hack?}{Tristan}{I agree, we should
  keep IT setups out of the picture. Depending on the internals of the
  science gateway, such hacks may or may not be feasible. I added a
  sentence to mention this possibility, without changing the score.}
Distributed engines are not available by default
(\texttt{S$_2$}=1). 
The scheduling of tasks on the infrastructure is
as complex as in any other architecture since the workflow engine
might implement any kind of scheduling policy (\texttt{S$_3$}=0).
\note{sil}{so same as the others = "easy"? i think this requires more explanation. maybe we should introduce a score for "dont know"?}

\paragraph{Specific features} Meta-workflows are not supported by default
(\texttt{O$_1$}=1).  
\note{sil}{why not? if the engine is ws-pgrade it can. i really think that meta-workflows is a property exclusively of the engine. you can see that in fig 2.V interaction *6}
Debugging is not an issue since the science
gateway can retrieve any information from the workflow engine directly
(\texttt{O$_2$}=0).

@@

\note{sil}{should we then also thank SURF foundation (for the dutchgrid), FP7 (for SCI_BUS, SHIWA and ER-flow), and COMMIT/ and HUvA HPCN for the funding for the amc science gateway activities. My participation has been minimal, and the amc gateways are not discussed in detail (my fault), so maybe it would be an overkill}
